\LARGE \bf Artifical Cognition for Social Human-Robot Interaction:\\ An Implementation

### 2016-06-14 paper.tex
Fixing language 5

Monitoring human activity is needed by the execution controllers to track the
engagement of the human and the progress of his/her actions. Full human action
and activity recognition is a task that requires knowledge and reasoning both on
high level facts like goals, intentions and plans, as well as bottom-up data
from agent and object motions. {\Spark} implements a set of simple temporal and
geometric heuristics on human hand trajectories and possible objects placements
to recognize simple atomic actions. Those primitive actions are assessed through
monitoring situations like ``an empty hand is close to an object on a table''
(precursor for a \emph{pick}), or ``a hand holding an object is over a
container'' (precursor for a \emph{throw}).  {\sc Spark} recognises a set of
such primitives that are relied upon by the execution controller to track the
state transitions in the human plans.

@@

The main weakness of our current implementation relates to the way semantics are
conveyed: because we use 2D markers for artifacts, we avoid issues related to object
segmentation and recognition. Each object has a unique identifier,
which enables us to load a proper CAD model suitable for 3D spatial reasoning
and prevent recognition mistakes in the knowledge base. While {\sc Spark}
algorithms themselves are ignorant of the input sources, and would work equally
well with a full object recognition stack, we do not have tackled it yet.

Besides, temporal reasoning (essential for accurate action recognition for
instance) is not properly addressed in our stack. Temporal reasoning is used
only locally, and does not allow for tracking of long sequences or global events.

@@

In our experience, natural language processing is one of the fields of
human-robot interaction for which the introduction of the semantic layer has
been most beneficial.  In~\cite{Lemaignan2011a} we explain the techniques and the
tool called {\sc Dialogs} that we have developed for natural English language
parsing and grounding, along with verbalisation and (minimalist) dialogue
management.

Natural language input (in experiments, we rely on an Android-based interface,
with Google speech recognition) is parsed into a grammatical structure, and
atoms of each sentence are resolved with the help of the ontology to ground
concepts like objects (\ie when a user says ``pick the can'', it resolves to which
instance of \emph{can} the user is referring to) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.
Figure~\ref{dialogs|ex} gives a example of the processing of a simple order.

\begin{figure}
    \centering
	\begin{tabular}{l|l}
	\emph{Initial human knowledge} &
	\emph{Human input}\\

@@

    	\stmt{book\_1 type Book} &
	``Give me the book.'' \\

    	\stmt{book\_1 hasColor blue} & \\
	\vspace{0.5em}\\
	\hline

	\emph{Generated partial statements} &
	\emph{Newly created statements}\\
	\hline
    	\stmt{?obj type Book} & 
	\stmt{human desires sit\_1} \\

	\hspace{0.2cm}$\Rightarrow$ \concept{?obj = book\_1}
    	& \stmt{sit\_1 type Give} \\
    	& \stmt{sit\_1 performedBy myself} \\
    	& \stmt{sit\_1 actsOnObject book\_01} \\
    	& \stmt{sit\_1 receivedBy human} \\
	\end{tabular}

    \caption{Processing of a simple, non-ambiguous order, taken
        from~\cite{Lemaignan2011a}. Thematic roles ({\tt performedBy}, {\tt
        actsOnObject}, {\tt receivedBy}) are automatically extracted by 
    {\sc Dialogs} from the order ``Give me the book.''.}

@@

The system supports quantification (``give me \{a | the | some | all | any |
...\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal} and added to the knowledge base
after checking for possible contradictions with existing knowledge),
interprets common temporal and place adverbs (like \emph{above} or
\emph{tomorrow}) and translates to a certain extent \emph{states} (``I'm
tired'') into \emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1
hasFeature tired}). We kindly refer the readers to~\cite{Lemaignan2011a} for a
complete discussion of these capabilities and the related algorithms.

@@

Because all components rely on the same RDF formalism to format their outputs,
the different communication modalities (\emph{explicit} like verbal, deictic or
based on gaze, or \emph{implicit} like postures) are presented in a homogeneous
way, as statements in the knowledge base. The dialogue grounding process makes
use of them at two distinct levels to provide seamless multi-modal interaction.

First, particular steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when several instances match
a category (the human says ``give me the bottle'' and the robot knows about
three bottles), the module may decide (it actually depends on the quantifier
preceding the class, see the example at section~\ref{moving-london}) to discard
some of them based on their \emph{visibility} for the speaker (implicit
communication context built on the human posture).

@@

Note that, while the system benefits from the complementary modalities, they are
not all required. The dialogue system can run with the verbal modality alone, at
the cost of a simpler interaction: if the human says ``this'' without the robot
tracking what the human points at, no \stmt{HUMAN pointsAt ...} fact is
available in the knowledge base, and the robot fallbacks on the anaphora
resolution step alone.

@@

The HATP \emph{planning domain} defines a set of methods describing how to
decompose a task and represents the procedural knowledge of the robot as well as
its knowledge about the actions that the human partner is able to achieve. It is
stored outside of the central knowledge base, using a specific formalism (see
the related discussion at the end of this section).

@@

The originality of HATP resides in its ability to produce plans for the robot's
actions \emph{as well as} for the other participants (humans or robots), that we
call \emph{shared plans}.

An important feature of HATP is that it treats agents as ``first-class
entities'' in the domain representation language. It can therefore
distinguish between the different agents in the domain as well as
between agents and the other entities such as tables and chairs. This
facilitates a post-processing step in HATP that splits the final
solution (sequence of actions) into two (or more if there are several
humans) synchronised solution
streams, one for the robot and one for the human, so that the streams may be executed in
parallel and synchronised when necessary (Figure~\ref{plan_hatp1}).

This effectively enriches the interaction capabilities
of the robot by allowing for a form of human behaviour \emph{prediction}, which
is used by the robot execution controller to monitor the engagement of the human
partner during joint tasks.

The planner also generates synchronisation points between the agents.  In the
plan depicted on Figure~\ref{plan_hatp1}, the robot needs to wait for the
success of the human agent's {\tt PUT} action (estimated through the fulfilment
of the action post-conditions, typically $\neg$\stmt{GREY\_TAPE isOn TABLE}) to
be able to start its own action.

@@

\caption{A plan produced by HATP with two action streams (human actions at the
top, robot actions at the bottom, {\sf PUTRV} stands here for {\it Put it so it is both
reachable and visible}). The arrow between the two streams represents
a synchronization point.}

@@

A cost and a duration function is associated to each action.  The duration
function provides a duration interval for the action achievement and is used, in
one hand, to schedule the different streams and, in the other hand, as an
additional cost function.

@@

Figure~\ref{plan_hatp2} illustrates such a socially-optimised plan where the
\emph{no wasted time} social rule is applied. The resulting shared plan is
considered better than the one depicted on Figure~\ref{plan_hatp1} according to
a global evaluation of the costs and rules.

@@

\caption{An alternative plan where the \emph{no wasted time} social rule is used to optimise the
total duration of the task. }

@@

Because HATP is a generic symbolic task planner, we have been able to design a
planning domain at a semantic level which is close to the one used in the
human-robot dialogue (the planner vocabulary contains concepts like
\texttt{give}, \texttt{table}, \texttt{is on}...). This leads to a natural
mapping (low ``impedance mismatch'') between the knowledge extracted from the situation
assessment or the dialogue, and the planner.

@@

\subsubsection{Desires and Experiences}

>>>>

Monitoring human activity is needed by the execution controllers to track the
engagement of the human and the progress of his/her actions. Full human action
and activity recognition is a task that requires knowledge and reasoning both on
high level facts like goals, intentions and plans, as well as bottom-up data
from agent and object motions. {\sc Spark} implements a set of simple temporal and
geometric heuristics on human hand trajectories and possible objects placements
to recognize simple atomic actions. Those primitive actions are assessed through
monitoring situations like ``an empty hand is close to an object on a table''
(precursor for a \emph{pick}), or ``a hand holding an object is over a
container'' (precursor for a \emph{throw}).  {\sc Spark} recognises a set of
such primitives that are relied upon by the execution controller to track the
state transitions in the human plans.

@@

In its current form, our situation assessment module makes two strong
assumptions: the objects are known in advance (hence, we can rely on proper 3D
CAD model for spatial reasoning) and the robot benefits of an almost perfect
perception, made possible by the use of fiducial markers.  Each object receives
a unique tag which enables us to accurately localize it in 3D and prevent
recognition ambiguities that would be then reflected in the knowledge base.
While {\sc Spark} algorithms themselves are ignorant of the input sources, and
would work equally well with a full object recognition stack, we are yet to
investigate this research area.

Besides, temporal reasoning (essential for accurate action recognition for
instance) is not properly addressed in our architecture. Temporal reasoning is used
only locally, and does not allow for tracking of long sequences or global events.

@@

Natural language is a basic interaction modality that we use in our system both
as input (processing of the human speech) and as output (verbalization of the
robot intents). Natural language processing is facilitated by our architecture
mostly processing informations with semantics that are already close to the
human-level. This section presents the main features of our speech processor,
{\sc Dialogs}, that include semantic and multi-modal grounding, and interactive
disambiguation. Algorithmic and implementation details are provided
in~\cite{Lemaignan2011a}.

We acquire natural speech input from the human participants through a custom
Android-based interface that relies on the Google speech recognition API for
speech-to-text and relays the textual transcript to the robot. It is parsed into
a grammatical structure (\emph{Part of Speech} tagging), and the atoms of each
sentence are resolved with the help
of the ontology to ground concepts like objects (\ie when a user says ``pick the
can'', it resolves to which instance of \emph{can} the user is referring to) and
actions.  Sentences are classified into questions, desires and statements, and
processed accordingly.  Figure~\ref{dialogs|ex} gives a example of the
processing of a simple, non-ambiguous command.

\begin{figure}
    \centering
    \begin{tabular}{l}
        \emph{Initial human knowledge} \\
        \hline
        \stmt{book\_1 type Book} \\
        \stmt{human\_1 rdf:type Human} \\
        ~\\
        ~\\
        ~\\
    \end{tabular}
    \begin{tabular}{l}
        \emph{Input}\\

@@

        \concept{human\_1} says:\\
        ``Give me the book'' \\
        ~\\
        ~\\
        ~\\

    \end{tabular}
    \begin{tabular}{l}

        \emph{Generated query} \\
        \hline
        \concept{find(?obj type Book)} \\ 
        \hspace{0.2cm}$\Rightarrow$ \concept{?obj = book\_1} \\
        ~\\
        ~\\
        ~\\
    \end{tabular}
    \begin{tabular}{l}

        \emph{Newly created statements}\\
        \hline
        \stmt{human\_1 desires sit\_1} \\
         \stmt{sit\_1 type Give} \\
         \stmt{sit\_1 performedBy myself} \\
         \stmt{sit\_1 actsOnObject book\_01} \\
         \stmt{sit\_1 receivedBy human} \\
    \end{tabular}

    \caption{Processing of a simple, non-ambiguous command, taken
        from~\cite{Lemaignan2011a}. Thematic roles ({\tt performedBy}, {\tt
        actsOnObject}, {\tt receivedBy}) are automatically extracted by 
    {\sc Dialogs} from the imperative sentence ``Give me the book.''.}

@@

The system supports quantification (``give me \{a | the | some | all | any |
...\} can''), thematic roles (action-specific predicates that qualify the
actions), interactive disambiguation (the robot asks questions when it needs
more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also supports knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal} and added to the knowledge base
after checking for possible contradictions with existing knowledge),
interprets common temporal and spatial adverbs (like \emph{above} or
\emph{tomorrow}) and translates simple expressions of internal state (``I'm
tired'') into \emph{experiences} (\stmt{human\_1 experiences state\_1, state\_1
hasFeature tired}, see also Section~\ref{sect:desires}). A full account of the
{\sc Dialogs} features and the corresponding algorithms is available
in~\cite{Lemaignan2011a}.

@@

Because the components of our architecture rely on the same RDF formalism to
represent their outputs, the different communication modalities (\emph{explicit}
like verbal, deictic or based on gaze, or \emph{implicit} like the position of
the human) are presented in a homogeneous way, as statements in the knowledge
base. The dialogue grounding process makes use of them at two distinct levels to
provide multi-modal concept grounding.

First, specific steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when several instances match
a category (the human says ``give me the bottle'' and the robot knows about
three bottles), the module may decide (it actually depends on the quantifier
preceding the class, see the example at section~\ref{moving-london}) to discard
some of them based on their \emph{visibility} for the speaker (implicit
communication context taking into account the human position).

@@

Note that, while the system benefits from the complementary modalities, they are
not all required. The dialogue system can run with the verbal modality alone, at
the cost of a simpler interaction: if the human says ``this'' without the robot
tracking what the human points at, no \stmt{human\_1 pointsAt ...} fact would
possibly be available in the knowledge base, and the robot fallbacks on the
anaphora resolution step alone.

@@

The HATP planning domain defines a set of methods describing how to
decompose a task and represents the procedural knowledge of the robot as well as
its knowledge about the actions that the human partner is able to achieve. It is
stored outside of the central knowledge base, using a specific formalism (see
the related discussion at the end of this section).

@@

The originality of HATP resides in its ability to produce plans for the robot's
actions as well as for the other participants (humans or robots), that we
call \emph{shared plans}.

HATP treats agents as ``first-class
entities'' in the domain representation language. It can therefore
distinguish between the different agents in the domain as well as
between agents and the other entities such as tables and chairs. This
enables a post-processing step that splits the final
solution (sequence of actions) into two (or more if there are several
humans) synchronised solution
streams, one for the robot and one for the human, so that the streams may be executed in
parallel and synchronised when necessary (Figure~\ref{plan_hatp1}).

This effectively enriches the interaction capabilities of the robot by providing
the system with what is in essence a prediction of the human behaviour. This
prediction is used by the robot execution controller to monitor the engagement
of the human partner during joint tasks.

The planner also generates synchronisation points between the agents. For
instance, in the plan depicted on Figure~\ref{plan_hatp1}, the human needs to
wait for the success of the robot's action {\tt PUTRV}. The robot monitors the
success of its own action (by checking for the fulfilment of the action
post-conditions; in this particular case \stmt{HUMAN sees BOTTLE} and
\stmt{HUMAN canReach BOTTLE}) to estimate what and when the human is likely to
perform his next action (here, \concept{TAKE(BOTTLE, TABLE)}). This, in turn,
allows the robot to monitor the human engagement and progress in the shared
plan.

@@

\caption{An example of plan produced by HATP for a task consisting in
    cooperatively moving objects into their associated bins. Two action streams
    are generated (human actions at the top, robot actions at the bottom, {\sf
    PUTRV} stands here for {\it Put it so it is both reachable and visible}).
    The arrow between the two streams represents a synchronization point.}

@@

A cost and a duration function are associated to each action.  The duration
function provides a duration interval for the action achievement and is used, in
one hand, to schedule the different streams and, in the other hand, as an
additional cost function.

@@

Figure~\ref{plan_hatp2} illustrates such a socially-optimised plan where the
\emph{no wasted time} social rule is applied: compared to the plan depicted in
Figure~\ref{plan_hatp1}, the robot first moves the bottle so that the human can
immediately take it and put it into the bin, thus removing the human idle time.
The resulting shared plan is considered better than the one depicted on
Figure~\ref{plan_hatp1} according to a global evaluation of the costs and rules.

@@

    \caption{An alternative plan for the task presented in
    Figure~\ref{plan_hatp1} where the \emph{no wasted time} social rule is used to optimise the
total duration of the task. }

@@

As HATP is a generic symbolic task planner and do not enforce any abstraction
level for the planning domain, we were able to design a domain built from
top-level tasks whose semantics are close to the one used in the human-robot
dialogue: the planner domain effectively contains concepts like \texttt{give},
\texttt{table}, \texttt{is on}. This leads to an effective mapping between the
knowledge extracted from the situation assessment or the dialogue, and the
planner.

@@

\subsubsection{Desires and Experiences}
\label{sect:desires}
